Lattices of acyclic pipe dreams

### 2023-03-09 acyclicFacets/draft/latticePipeDreams.tex
missing pictures

As already mentioned, the increasing flip poset on reversing pipe dreams is isomorphic to the Tamari lattice, which is a lattice quotient of the weak order.
In this section, we extend this result to any permutation~$\omega$ by showing that the sets of linear extensions of pipe dreams of~$\acyclicPipeDreams(\omega)$ partition the interval~$[e,w]$ (\cref{subsec:linearExtensions}), that this partition actually defines a lattice congruence of the weak order on~$[e,w]$ (\cref{subsec:pipeDreamCongruence}), and that the Hasse diagram of the quotient by this congruence is the increasing flip graph on acyclic pipe dreams of~$\acyclicPipeDreams(\omega)$ (\cref{subsec:pipeDreamQuotient}).
This section goes straight to the proof of this property, and leaves alternative perspectives on this quotient to \cref{sec:furtherTopics}.
\vincent{add picture}

@@

\begin{figure}[ht]
	\centerline{\includegraphics[scale=1.3]{counterExampleRestrictionIncreasingFlipPoset}}
	\caption{Two acyclic pipe dreams~$\acyclicPipeDreams(126543)$ connected by a sequence of increasing flips in~$\pipeDreams(126543)$ but by no sequence of increasing flips in~$\acyclicPipeDreams(126543)$.}
	\label{fig:counterExampleRestrictionIncreasingFlipPoset}
\end{figure}
\end{remark}

@@

\begin{remark}
\label{rem:linearExtensionsPartitionSubwordComplexA}
Before proving this theorem, we would like to make the following remarks about items~\eqref{item:convex} and~\eqref{item:cover}.
\begin{enumerate}
\setcounter{enumi}{0}
\item There are cases where $[e,\omega] \ne \linearExtensions(Q,\omega)$. \cesarm{add discussion}
%
\begin{figure}[h]
	\includegraphics[width=0.52\textwidth]{interval_not_contained1} \quad
	\includegraphics[width=0.43\textwidth]{interval_not_contained2}
	\caption{A facet of a subword complex~$\subwordComplex(Q, \omega)$ whose linear extensions all lie outside the interval~$[e, \omega]$.}
	\label{fig:outOfInterval}
\end{figure}
%
\setcounter{enumi}{2}
\item $\linearExtensions(I)$ is not an interval in general. See for instance the example~\cite[Figure~9]{PilaudStump-brickPolytope}. 
In that figure, the elements of the week order $\mathfrak{S}_4$ that are linear extensions of the same acyclic facet of a given subword complex are grouped together. Some of the groups in the figure are not intervals.  \\
\end{enumerate}
\end{remark}

>>>>

As already mentioned, the increasing flip poset on reversing pipe dreams is isomorphic to the Tamari lattice, which is a lattice quotient of the weak order.
In this section, we extend this result to any permutation~$\omega$ by showing that the sets of linear extensions of pipe dreams of~$\acyclicPipeDreams(\omega)$ partition the interval~$[e,w]$ (\cref{subsec:linearExtensions}), that this partition actually defines a lattice congruence of the weak order on~$[e,w]$ (\cref{subsec:pipeDreamCongruence}), and that the Hasse diagram of the quotient by this congruence is the increasing flip graph on acyclic pipe dreams of~$\acyclicPipeDreams(\omega)$ (\cref{subsec:pipeDreamQuotient}).
This section goes straight to the proof of this property, and leaves alternative perspectives on this quotient to \cref{sec:furtherTopics}.

@@

\begin{figure}[ht]
	\centerline{\includegraphics[scale=1.2]{counterExampleRestrictionIncreasingFlipPoset}}
	\caption{Two acyclic pipe dreams~$\acyclicPipeDreams(126543)$ connected by a sequence of increasing flips in~$\pipeDreams(126543)$ but by no sequence of increasing flips in~$\acyclicPipeDreams(126543)$.}
	\label{fig:counterExampleRestrictionIncreasingFlipPoset}
\end{figure}
\end{remark}

@@

\begin{remark}
\label{rem:linearExtensionsPartitionSubwordComplexA}
Before proving this theorem, we would like to make the following remarks about items~\eqref{item:convex} and~\eqref{item:cover}:
\begin{enumerate}
\setcounter{enumi}{0}
\item There are cases where $[e,\omega] \ne \linearExtensions(Q,\omega)$. Namely, some of the sets~$\linearExtensions(I)$ may not be included in~$[e,\omega]$, and may even not meet~$[e,\omega]$. See \cref{fig:counterExample1}.
%
%\begin{figure}[h]
%	\includegraphics[width=0.52\textwidth]{interval_not_contained1} \quad
%	\includegraphics[width=0.43\textwidth]{interval_not_contained2}
%	\caption{A facet of a subword complex~$\subwordComplex(Q, \omega)$ whose linear extensions all lie outside the interval~$[e, \omega]$.}
%	\label{fig:outOfInterval}
%\end{figure}
%
\begin{figure}[p]
	\centerline{\includegraphics[scale=.8]{counterExample1}}
	\caption{The sets~$\linearExtensions(I)$ (blue bubbles) are not always subsets of, and may even not meet~$[e,\omega]$ (green bubble).}
	\label{fig:counterExample1}
\end{figure}

\setcounter{enumi}{2}
\item Some of the sets~$\linearExtensions(I)$ may not be intervals. See \cref{fig:counterExample2}. Another example with~$\omega = \omega_\circ$ can be found in~\cite[Figure~9]{PilaudStump-brickPolytope}. 
%
\begin{figure}[p]
	\centerline{\includegraphics[scale=.8]{counterExample2}}
	\caption{The sets~$\linearExtensions(I)$ (blue bulles) are not always intervals.}
	\label{fig:counterExample2}
\end{figure}
\end{enumerate}
\cref{fig:counterExample1,fig:counterExample2} represent type~$A$ subword complexes.
As in~\cite{PilaudPocchiola, PilaudSantos-brickPolytope, PilaudStump-brickPolytope}, we thus use pseudoline arrangements on sorting networks to represent their facets (and we have represented their contact graphs together).
\end{remark}
